Towards a Mechanized Proof of Noninterference

### 2013-12-18 paper/paper.tex
Very short and sweet abstract.

Abstract goes here, summarize what this paper covers and what we ended up with.

@@

\begin{figure}
  \begin{align}
    v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag{values} \\
    a & ::= v \tag{answers} \\
    e & ::= a \: | \: v v \: | \: \textsf{if } v \: e_1 \: e_2 \: | \: \textsf{let
    } x := v \textsf{ in } e \tag{expressions} \\
  \end{align}
  \caption{Syntax of \langName}
  \label{fig:syntax}
\end{figure}

@@

Outline the operational semantics of the baby language we developed.  Outline
any interesting similarities or differences between us and Pottier and Simonet.

\begin{figure}
  Basic reductions
  \begin{align}
    (\textsf{fix} f . \lambda x . e) v
    \: \to \: &
    e[x \Leftarrow v][f \Leftarrow \textsf{fix} f . \lambda x . e]
    \tag{$\beta$}
    \label{eq:beta}
    \\
    \textsf{if } 1 \: e_1 \textsf { } e_2
    \: \to \: &
    e_1
    \tag{if-true}
    \label{eq:if-true}
    \\
    \textsf{if } (k \neq 1) \textsf { } e_1 \textsf { } e_2
    \: \to \: &
    e_2
    \tag{if-false}
    \label{eq:if-false}
    \\
    \textsf{let }  x := v \textsf{ in } e
    \: \to \: &
    e[x \Leftarrow v]
    \tag{let}
    \label{eq:let}
  \end{align}

@@

@@

@@

\begin{lemma}[Completeness]
  \label{lem:completeness}
  Assume $\lift{e}_i \to^* a_i$ for all $i \in \{1, 2\}$.  Then, there exists an
  answer $a$ such that $e \to^* a$.
\end{lemma}
\begin{proof}
  TODO
\end{proof}

@@

\begin{lemma}
  Let $H$ be an upward-closed subset of the label lattice $\mathcal{L}$.  Let $l
  \not\in H$.  If, using $H$ for the set of high labels, $\typeRule{
  }{e}{\textsf{int}^l}$ and $e \to^* v$ then $\lift{v}_1 = \lift{v}_2$.
\end{lemma}
\begin{proof}
  TODO
\end{proof}

\begin{thm}[Noninterference]
  Choose $l, h \in \mathcal{L}$ such that $h \not\leq l$.  Let $h \guards t$.
  Assume $\typeRule{(x \mapsto t)}{e}{\textsf{int}^l}$, where $e$is a \langName
  expression.  If, for all $i \in \{1, 2\}$, $\typeRule{ }{v_i}{t}$ and $e[x
  \Leftarrow v_i] \to^* v_i'$ hold, then $v_1' = v_2'$.
\end{thm}
\begin{proof}
  TODO
\end{proof}

>>>>

In this paper we discuss our experience trying to mechanize a proof of
noninterference for a simple language with security labels on the types.
Although we did not complete the full proof because we ran out of time, we were
able to find small inaccuracies and subtleties in the proofs from the paper we
based our proof on.  We discuss what would be involved in completing the proof
and how this relates to other work.

@@

Our toy language for proving noninterference, \langName, is a simplified version
of \origLang from Simonet and Pottier's paper~\cite{InfoFlowML}.  Since we had a
limited amount of time to work on this project, we limited ourselves to a subset
of the language that drops the following features:

\begin{itemize}
  \item References and associated memory operations.
  \item Paired values.
  \item Exceptions.
  \item Sequencing operations.
\end{itemize}

While the language becomes much much less interesting for actual programming, we
feel that these features do not fundamentally change the problem of proving
noninterference.  In order to mechanize a proof of noninterference for a
language closer to \origLang, it would require simply adding mechanized proofs
for the additional cases where appropriate.

\textbf{TODO: Talk about what the bracketed expressions and values are.}

Furthermore, primarily due to a matter of taste, we decided to replace the
\textsf{case} construct from \origLang for an \textsf{if} construct, which
evaluates the first expression if the test evaluates to 1 and evaluates to the
second expression otherwise.

Through these changes, our language has the syntax shown in
Figure~\ref{fig:syntax}.  As shown, final states, answers, in the language are
now simple values because there are no exceptions.

\begin{figure}[h!]
  \begin{align}
    v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag{values} \: |
    \: <v | v> \\
    a & ::= v \tag{answers} \: | \: <a | a> \\
    e & ::= a \: | \: v v \: | \: \textsf{if } v \: e \: e \: | \: \textsf{let
    } x := v \textsf{ in } e \tag{expressions} \: | \: <e | e> \\
  \end{align}
  \caption{Syntax of \langName}
  \label{fig:syntax}
\end{figure}

In the following two subsections, we introduce the operational semantics and
type system for the language.

@@

\langName has an operational semantics that should be familiar to the reader who
has seen a single step semantics for a basic lambda calculus.  The interesting
additions are the same as those in \origLang, where there are additional rules
for handling the bracketed values.

\textbf{TODO: Talk about what the rules do at a high level.}

\begin{figure}[h!]
  Basic reductions
  \begin{align}
    (\textsf{fix} f . \lambda x . e) v
    \: \to \: &
    e[x \Leftarrow v][f \Leftarrow \textsf{fix} f . \lambda x . e]
    \tag{$\beta$}
    \label{eq:beta}
    \\
    \textsf{if } 1 \: e_1 \textsf { } e_2
    \: \to \: &
    e_1
    \tag{if-true}
    \label{eq:if-true}
    \\
    \textsf{if } (k \neq 1) \textsf { } e_1 \textsf { } e_2
    \: \to \: &
    e_2
    \tag{if-false}
    \label{eq:if-false}
    \\
    \textsf{let }  x := v \textsf{ in } e
    \: \to \: &
    e[x \Leftarrow v]
    \tag{let}
    \label{eq:let}
  \end{align}

@@

\textbf{TODO: Talk more about the rules and what they're doing.  Specifically,
discuss what the heck H and pc are doing.}

@@

\textbf{TODO: Add in text between lemmas and theorems discussing what each one
is at a high level.}

@@

\begin{lemma}[Completeness]
  \label{lem:completeness}
  Assume $\lift{e}_i \to^* a_i$ for all $i \in \{1, 2\}$.  Then, there exists an
  answer $a$ such that $e \to^* a$.
\end{lemma}
\begin{proof}
  \textbf{TODO}
\end{proof}

@@

\begin{lemma}
  Let $H$ be an upward-closed subset of the label lattice $\mathcal{L}$.  Let $l
  \not\in H$.  If, using $H$ for the set of high labels, $\typeRule{
  }{e}{\textsf{int}^l}$ and $e \to^* v$ then $\lift{v}_1 = \lift{v}_2$.
\end{lemma}
\begin{proof}
  \textbf{TODO}
\end{proof}

\begin{thm}[Noninterference]
  Choose $l, h \in \mathcal{L}$ such that $h \not\leq l$.  Let $h \guards t$.
  Assume $\typeRule{(x \mapsto t)}{e}{\textsf{int}^l}$, where $e$is a \langName
  expression.  If, for all $i \in \{1, 2\}$, $\typeRule{ }{v_i}{t}$ and $e[x
  \Leftarrow v_i] \to^* v_i'$ hold, then $v_1' = v_2'$.
\end{thm}
\begin{proof}
  \textbf{TODO}
\end{proof}
